---
title: "Dewlap Analysis"
author: "Lauren E Johnson"
date: "11/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggpubr)
```


#### **Background**  
**Response Variable** = `dewlappermin`  
**Predictor Variables** = `dewsize` and `visibility`  
**Covariate** = `svl`  
**Other variables** = `pushbobpermin` and `movepermin`  

*Question* Do dewlap display behaviors of male anoles align with principles of signal efficacy?  

#### **Data Analysis**  


**Load Data**  
Also create separate dataframes for each species to make analysis later a little easier 

```{r}

lat <- c(18.4582)
lon <- c(-77.3985)
id <- c('Discovery Bay Marine Laboratory')
maps::map(database = 'world', region = 'jamaica', col= 'gray90', fill=TRUE)
points(x = lon, y = lat, pch=19, col="red", cex=2, )
text(lon, lat, id, pos = 4, col = 'blue', cex = 0.5)


```


```{r load data}
load(here::here('data/jm_anoles.Rda'))

#Creating separate dataframes for each species
g <- jmanoles[which(jmanoles$species == "grahami"),]
l <- jmanoles[which(jmanoles$species == "lineatopus"),]
v <- jmanoles[which(jmanoles$species == "valencienni"),]










# knitr::kable(head(df))
```


**Meeting with Don**

Meeting wit Don...
Plot original data but then report that the stats were conducted with it log transformed 
try 3d surface plot to see how it turns out
Mosimann size adjustment - allows pulling out of variation in trait that is masked by correlation with size (regressions can't handle collinearity)

analyses shows that after you remove the effect of body size, get a very strong effect of dewlap size (even greater than dewlap size by itself and body size bby itsn't)
large male with small dewlap 
small male with large dewlap 

exaggerates difference in dewlap size
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4896994/
```{r}
# 2 methods for extracting svl from dewlap size - (1) residuals of a linear regression of svl and dewlap size & (2) Mosimann adjustment, which is the log10(dewsize) - log10(svl)

# (1) residuals 
# creating a regression of svl and dewlap size for grahami
g.svdew.lm <- lm(log10(dewsizemax) ~ log10(svl), data = g)
plot(g.svdew.lm)
# save the residuals into a vector for analysis later
g.res.dew <- g.svdew.lm$residuals
# add the residuals to the dataframe for grahami 
g$res_dewsize <- g.res.dew

# do the same thing for lineatopus
l.svdew.lm <- lm(log10(dewsizemax) ~ log10(svl), data = l)
plot(l.svdew.lm)
l.res.dew <- l.svdew.lm$residuals
l$res_dewsize <- l.res.dew

# (2) Mosimann adjustment 
# grahami 
g.mos_dewsize <- log10(g$dewsizemax) - log10(g$svl)
g$mos_dewsize <- g.mos_dewsize
# lineatopus
l.mos_dewsize <- log10(l$dewsizemax) - log10(l$svl)
l$mos_dewsize <- l.mos_dewsize

df <- rbind(g, l)

#Add dewlap and pushbob rates per hour since they will be bigger numbers and not fractions
df$dewlapperhour <- 60*(df$dewlappermin)
df$pushbobperhour <- 60*(df$pushbobpermin)


# add the log of svl and dewlap size 
df$logsvl <- log10(df$svl)
df$logdewsizemax <- log10(df$dewsizemax)
df$logdewlappermin <- log10(df$dewlappermin + 1)
df$logvisibility <- log10(df$visibility)

#Remove individual without NAs for dewlap behaviors

df <- df[!is.na(df$dewlapperhour),]

# recreate g and l dataframes

g <- df[which(df$species=='grahami'),]

l <- df[which(df$species=='lineatopus'),]
```


12/14/2022
```{r}

#Non of the habitat variables are correlated with each other

g.hab <- data.frame(height = g$height, diamter = g$diameter, visibility = g$visibility)

g.hab.cor <- cor(g.hab, method = "pearson")

library("Hmisc")
res2 <- rcorr(as.matrix(g.hab))
res2

## do the same thing for the behavior data
g.cor <- data.frame(svl = g$svl, dewarea = g$dewsizemax, height = g$height, diamter = g$diameter, visibility = g$visibility, move = g$movepermin, pushbobs = g$pushbobpermin, dewexten = g$dewlappermin)

g.cor.p <- cor(g.cor, method = "pearson")

library("Hmisc")
res1 <- rcorr(as.matrix(g.cor), type = "pearson")
res1

library(corrplot)
corrplot(g.cor.p, type = "upper", sig.level = 0.05, insig = "blank")

#lineatopus 
l.cor <- data.frame(svl = l$svl, dewarea = l$dewsizemax, height = l$height, diamter = l$diameter, visibility = l$visibility, move = l$movepermin, pushbobs = l$pushbobpermin, dewexten = l$dewlappermin)

res2 <- rcorr(as.matrix(l.cor), type = "pearson")
res2


```










Are the visibility of habitats different between grahami (trunk-crown) and lineatopus (trunk-ground) habitats   
```{r}

ggdensity(df, x = 'visibility', add = 'mean', rug = T, color = 'species', fill = 'species')

t.test(visibility~species, data = df)

# log transforming visibility does nothing, so don't need to worry about it for analyses
ggdensity(df, x = 'logvisibility', add = 'mean', rug = T, color = 'species', fill = 'species')

gghistogram(df, x = 'visibility', add = 'mean', rug = T, color = 'species', fill = 'species')

ggdensity(df, x = 'height', add = 'mean', rug = T, color = 'species', fill = 'species')


gghistogram(df, x = 'height', add = 'mean', rug = T, color = 'species', fill = 'species')

```

What is the variance in dewlap size and svl across grahami and lineatopus?
```{r}

d <- ggdensity(df, x = 'dewsizemax', add = 'mean', rug = T, color = 'species', fill = 'species')
s <- ggdensity(df, x = 'svl', add = 'mean', rug = T, color = 'species', fill = 'species')

ggarrange(d, s, heights = c(4, 4),
          ncol = 1, nrow = 2, align = "v")

d <- ggdensity(df, x = 'logdewsizemax', add = 'mean', rug = T, color = 'species', fill = 'species')
s <- ggdensity(df, x = 'logsvl', add = 'mean', rug = T, color = 'species', fill = 'species')

ggarrange(d, s, heights = c(4, 4),
          ncol = 1, nrow = 2, align = "v")

d <- gghistogram(df, x = 'dewsizemax', add = 'mean', rug = T, color = 'species', fill = 'species')
s <- gghistogram(df, x = 'svl', add = 'mean', rug = T, color = 'species', fill = 'species')

ggarrange(d, s, heights = c(4, 4),
          ncol = 1, nrow = 2, align = "v")
```

```{r}
ggplot(df, aes(svl, dewsizemax, color = species)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

```


How does dewlap per hour vary across species?

```{r}
ggdensity(df, x = 'dewlapperhour', add = 'mean', rug = T, color = 'species', fill = 'species')

ggdensity(df, x = 'dewlappermin', add = 'mean', rug = T, color = 'species', fill = 'species')

ggdensity(df, x = 'logdewlappermin', add = 'mean', rug = T, color = 'species', fill = 'species')

gghistogram(df, x = 'dewlapperhour', add = 'mean', rug = T, color = 'species', fill = 'species')

gghistogram(df, x = 'pushbobperhour', add = 'mean', rug = T, color = 'species', fill = 'species')

```


How does dewlap size vary with svl between grahami and lineatopus?
```{r}
ggscatter(df, x = 'svl', y = 'dewsizemax', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'svl', y = 'res_dewsize', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logsvl', y = 'res_dewsize', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'svl', y = 'mos_dewsize', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logsvl', y = 'logdewsizemax', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'dewsizemax', y = 'res_dewsize', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logsvl', y = 'dewlappermin', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')
```

******R2 and P value for the relationship between SVL and Dewsize Max******* 
This has all the analyses I use for the powerpoint presentation 
```{r}
g.lm <- lm(log(svl)~log(dewsizemax), data = g)

summary(g.lm)
# R2 = 0.88
# p < 0.001

l.lm <- lm(log(svl)~log(dewsizemax), data = l)

summary(l.lm)
# R2 = 0.68
# p < 0.001

# outputs are put on sicb graphs

summary(lm(logdewlappermin~visibility, data = g))
summary(lm(logdewlappermin~res_dewsize, data = g))
summary(lm(logdewlappermin~logsvl, data = g))
summary(lm(logdewlappermin~logdewsizemax, data = g))
summary(lm(visibility~logsvl, data = g))
summary(lm(visibility~logdewsizemax, data = g))

summary(lm(logdewlappermin~visibility, data = l))
summary(lm(logdewlappermin~res_dewsize, data = l))
summary(lm(logdewlappermin~logsvl, data = l))
summary(lm(logdewlappermin~logdewsizemax, data = l))
summary(lm(visibility~logsvl, data = l))
summary(lm(visibility~logdewsizemax, data = l))


# residual dewsize and svl and visibility 
g.lm.res <- lm(logdewlappermin~res_dewsize + logsvl + visibility, data = g, na.action = na.omit)
summary(g.lm.res)
car::Anova(g.lm.res)

l.lm.res <- lm(logdewlappermin~res_dewsize + logsvl + visibility, data = l, na.action = na.omit)
summary(l.lm.res)
car::Anova(l.lm.res)

# absolute dewsize and visibility
g.lm.abs <- lm(logdewlappermin~logdewsizemax + visibility, data = g, na.action = na.omit)
g.lm.abs.sum <- summary(g.lm.full)
car::Anova(g.lm.full)

l.lm.abs <- lm(logdewlappermin~logdewsizemax + visibility, data = l, na.action = na.omit)
summary(l.lm.full)
car::Anova(l.lm.full)

library(knitr)
t1 <- broom::tidy(car::Anova(g.lm.res))
t2 <- broom::tidy(car::Anova(l.lm.res))
t3 <- broom::tidy(car::Anova(g.lm.abs))
t4 <- broom::tidy(car::Anova(l.lm.abs))

library(gt)
# Resuls of ANOVA on linear regression analysis
t1f<- gt(t1) %>%
    cols_label(term = 'Source', sumsq = 'SS', statistic = 'F', p.value = 'p-value') %>%
    fmt_number(columns = c('sumsq', 'statistic', 'p.value'), decimals = 4) %>%
    fmt_missing(columns = 2:5, missing_text = " ") %>% tab_header(title = 'A. grahami')

t2f <- gt(t2) %>%
    cols_label(term = 'Source', sumsq = 'SS', statistic = 'F', p.value = 'p-value') %>%
    fmt_number(columns = c('sumsq', 'statistic', 'p.value'), decimals = 4) %>%
    fmt_missing(columns = 2:5, missing_text = " ") %>% tab_header(title = 'A. lineatopus')

t3f <- gt(t3) %>%
    cols_label(term = 'Source', sumsq = 'SS', statistic = 'F', p.value = 'p-value') %>%
    fmt_number(columns = c('sumsq', 'statistic', 'p.value'), decimals = 4) %>%
    fmt_missing(columns = 2:5, missing_text = " ") %>% tab_header(title = 'A. grahami')

t4f <- gt(t4) %>%
    cols_label(term = 'Source', sumsq = 'SS', statistic = 'F', p.value = 'p-value') %>%
    fmt_number(columns = c('sumsq', 'statistic', 'p.value'), decimals = 4) %>%
    fmt_missing(columns = 2:5, missing_text = " ") %>% tab_header(title = 'A. lineatopus')

gtsave(t1f, filename = 'g.lm.res.rtf', exapnd = 10)
gtsave(t2f, filename = 'l.lm.res.rtf', exapnd = 10)
gtsave(t3f, filename = 'g.lm.abs.rtf', exapnd = 10)
gtsave(t4f, filename = 'l.lm.abs.rtf', exapnd = 10)

```
********** plots for sicb talk some are in repeatability_dewlap_anlaysis****************
```{r}
ggplot(g, aes(res_dewsize, dewlappermin)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  #scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

ggplot(l, aes(res_dewsize, dewlappermin)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  #scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

gghistogram(df, x = 'svl', add = 'mean', rug = T,  fill = 'species', bins = 40, binwidth = 0.5, add_density = T, xlim = c(40,70)) + 
  facet_wrap(~species, dir = 'v') 


ggplot(g, aes(svl, visibility)) +
  geom_point() + 
  theme_classic() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

ggplot(l, aes(svl, visibility)) +
  geom_point() + 
  theme_classic() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)
```






Are larger lizards found in more visible habitats? But SVL is a wide range, so really easy to get a correlation I bet!!
```{r}
ggscatter(df, x = 'logsvl', y = 'visibility', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logdewsizemax', y = 'visibility', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'res_dewsize', y = 'visibility', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'mos_dewsize', y = 'visibility', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logdewlappermin', y = 'visibility', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

```
are larger individuals found at heigher perch heights?

```{r}
ggscatter(df, x = 'logsvl', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logdewsizemax', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'res_dewsize', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'mos_dewsize', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(df, x = 'logdewlappermin', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

# just grahami
ggscatter(g, x = 'height', y = 'dewlappermin', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species), method = 'pearson')
ggscatter(g, x = 'height', y = 'dewsizemax', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species), method = 'pearson')
ggscatter(g, x = 'logsvl', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species), method = 'pearson')

ggscatter(l, x = 'logsvl', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species), method = 'pearson')
ggscatter(l, x = 'dewsizemax', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species), method = 'pearson')
ggscatter(l, x = 'dewlappermin', y = 'height', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species), method = 'pearson')


# are highter perches within each species more visible?

ggscatter(df, x = 'height', y = 'visibility', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')
# NO
```

**Correlation Matrix**
Creating a correlation matrix that compares the correlation among `svl`, `dewsizemax`, `visibility`, `dewlappermin`, `pushbobpermin`, and `movepermin` for all the lizard species together and each species separately.  

Nubmers in the matrix represent the correlation coefficient. The color and size of the circle represnet the magnitude and direciton of the correlation. And an `x` over the correlation coefficient signifies that it is not a statistically signficant correlation. 
```{r correlation matrix}
# correlation matrix function 

crmtrx <- function(df, speciesname){
  
  # subsetting the continuous variables for the matrix
  cordata <- df[,c('svl', 'dewsizemax', 'visibility', 'dewlappermin', 'pushbobpermin', 'movepermin')]
  
  M <- cor(cordata, method = "pearson", use = "complete.obs")

  testRes <- corrplot::cor.mtest(cordata, conf.level = 0.95)

  corrplot::corrplot(M, p.mat = testRes$p, method = 'circle', type = 'lower',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE, title = speciesname)
}


crmtrx(jmanoles, 'All')
crmtrx(g, 'grahami')
crmtrx(l, 'lineatopus')
crmtrx(v, 'valencienni')


```

For grahami, dewlappermin was strongly correlated with visibility and pushbobspermin
For lineatopus, dewlappermin was moderately correlated with dewlap size and svl 

**Regression Models**

Before continuing, for valencienni, we have both males and females. All the regression analyses have males and females clumped together, but we probably can't do that because when running t tests there are differences in dewlap behavior, dewlap size, and svl between the sexes (however, there is not a difference in the visibility of the habitats they were found displaying at).

```{r}

t.test(data = v, dewlappermin ~ sex)

t.test(data = v, dewsizemax ~ sex)

t.test(data = v, svl ~ sex)


t.test(data = v, visibility ~ sex)

```
There is a significant different between the sexes in dewlappermin, dewsize, and svl. However, there is no difference between the sexes in the visibility of the perches they chose. 


What is the distribution of the size adjusted dewlap size values
```{r}


gghistogram(df, x = 'res_dewsize', add = 'mean', rug = T, color = 'species', fill = 'species')

gghistogram(df, x = 'mos_dewsize', add = 'mean', rug = T, color = 'species', fill = 'species')


ggdensity(df, x = 'res_dewsize', add = 'mean', rug = T, color = 'species', fill = 'species')

ggdensity(df, x = 'mos_dewsize', add = 'mean', rug = T, color = 'species', fill = 'species')


```



Now that we have the size adjusted values, lets run the regression models to test whether visibility or dewlap size predicts dewlap dispaly behavior 
```{r}

ggscatter(g, x = 'mos_dewsize', y = 'dewlappermin', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(g, x = 'res_dewsize', y = 'dewlappermin', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(l, x = 'mos_dewsize', y = 'dewlappermin', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

ggscatter(l, x = 'res_dewsize', y = 'dewlappermin', color = 'species', add = 'reg.line', conf.int = T, conf.int.level = .95) + stat_cor(aes(color = species),method = 'pearson')

# Linear Regression 
dew.lm <- lm(log(dewlappermin+1) ~ visibility + res_dewsize, data = g)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)


dew.lm <- lm(log(dewlappermin+1) ~ visibility + mos_dewsize, data = g)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)

# is there interactions

dew.lm <- lm(log(dewlappermin+1) ~ visibility * res_dewsize, data = g)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)


dew.lm <- lm(log(dewlappermin+1) ~ visibility * mos_dewsize, data = g)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)

##

dew.lm <- lm(log(dewlappermin+1) ~ visibility + res_dewsize, data = l)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)


dew.lm <- lm(log(dewlappermin+1) ~ visibility + mos_dewsize, data = l)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)

# is there interacitons

dew.lm <- lm(log(dewlappermin+1) ~ visibility * res_dewsize, data = l)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)


dew.lm <- lm(log(dewlappermin+1) ~ visibility * mos_dewsize, data = l)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)


dew.lm <- lm(log(dewlappermin+1) ~ mos_dewsize + svl, data = l)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)

dew.lm <- lm(log(dewlappermin+1) ~ dewsizemax + visibility, data = g)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)
# # Residuals 
# dewlap.lm <- lm(log(dewlappermin+1) ~ visibility + res_dew, data = subset(jmanoles, species == 'grahami'))
# 
# plot(dewlap.lm)
# summary(dewlap.lm)
# car::Anova(dewlap.lm)
# 
# # Mosimann size adjustment
# 
# log_dewsize <- log10(g$dewsizemax) - log10(g$svl) # relative dewlap size is signicant dewlap size adjusted for SVL
# 
# dewlap.lm <- lm(log(dewlappermin+1) ~ visibility + log_dewsize , data = subset(jmanoles, species == 'grahami'))
# 
# plot(dewlap.lm)
# 
# summary(dewlap.lm)
# # want the min and max residuals symmetric (what the QQplot is showing)
# # estimat/std error = t value
# 
# car::Anova(dewlap.lm) # type II automaticall hols the other value constant will calculating second 
# 
# anova(dewlap.lm) # type 1 heirarchical variation explained 

# when model is significant summary(model) pvalue and f statistic but the individual values aren't 

```
Trying a scatter plot grouped by visibility color


More cleaned versions of the anlayses I want to run
```{r}

# absolute dewlap size


dew.lm <- lm(log(dewlappermin+1) ~ visibility * dewsizemax, data = g)
plot(dew.lm)
summary(dew.lm) # model itself is signficant, but not individual variables
car::Anova(dew.lm) # visibility is significant, dewsizemax trending

dew.lm <- lm(log(dewlappermin+1) ~ visibility * dewsizemax, data = l)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 

 
# no interaction, so took it out of model 
dew.lm <- lm(log(dewlappermin+1) ~ visibility + res_dewsize + logsvl, data = g)
plot(dew.lm)
summary(dew.lm) # model itself is signficant, but not individual variables
car::Anova(dew.lm) # visibility is significant, dewsizemax trending

dew.lm <- lm(log(dewlappermin+1) ~ visibility + res_dewsize + logsvl, data = l)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 


dew.lm <- lm(log(dewlappermin+1) ~ visibility, data = g)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 

####
dew.lm <- lm(log(dewlappermin+1) ~ visibility * dewsizemax, data = g)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 


dew.lm <- lm(log(dewlappermin+1) ~ visibility * res_dewsize, data = g)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 



dew.lm <- lm(log(dewlappermin+1) ~ visibility * dewsizemax, data = l)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 

dew.lm <- lm(log(dewlappermin+1) ~ visibility + dewsizemax, data = l)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm)

dew.lm <- lm(log(dewlappermin+1) ~ visibility + res_dewsize, data = l)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 

```



```{r}

library(ggplot2)

# showing visibility with color and size, and dewsize on the x axis
ggplot(g, aes(mos_dewsize, dewlappermin, color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() + 
  geom_smooth(method = 'lm', se = FALSE, color = 'black')

ggplot(g, aes(res_dewsize, dewlappermin, color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() + 
  geom_smooth(method = 'lm', se = FALSE, color = 'black')

ggplot(g, aes(dewsizemax, dewlappermin, color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() + 
  geom_smooth(method = 'lm', se = FALSE, color = 'black')


ggplot(g, aes(svl, dewlappermin, color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() + 
  geom_smooth(method = lm, se = FALSE)
##

ggplot(g, aes(mos_dewsize, log(dewlappermin+1), color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() + 
  geom_smooth(method = 'lm', se = FALSE, color = 'black')

# showing dewsize with color and size, and visibility on the x axis
ggplot(g, aes(visibility, dewlappermin, color = dewsizemax)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = dewsizemax), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black') + ylim(0, 3.2)

ggplot(g, aes(visibility, dewlappermin, color = res_dewsize)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = res_dewsize), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black') + ylim(0, 3.2)

ggplot(g, aes(visibility, log(dewlappermin+1), color = mos_dewsize)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = mos_dewsize), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black')

ggplot(g, aes(visibility, dewlappermin, color = mos_dewsize)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = mos_dewsize), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black') + ylim(0, 3.2)
# geom_smooth method  = 'lm'
  
```

Plot graphs showig the effect of mos_dew size on dewlapppermin for lineatopus

```{r}
# keeping with theme for lin 
ggplot(l, aes(mos_dewsize, log(dewlappermin+1), color = mos_dewsize)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = mos_dewsize), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black')


# geom_smooth method  = 'lm'

ggplot(l, aes(dewsizemax, dewlappermin, color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black')

ggplot(l, aes(res_dewsize, dewlappermin, color = visibility)) +
  scale_color_gradient(low = "yellow", high = "red") +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = visibility), alpha = 0.75) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black')

# no keeping with theme for grahami
ggplot(l, aes(dewsizemax, dewlappermin)) +
  geom_point()+
  #geom_point(aes(colour = visibility)) + 
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black')
# geom_smooth method  = 'lm'

```

Final SICB graphs

```{r}
# grahami
ggplot(g, aes(visibility, dewlappermin)) +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = dewsizemax), alpha = 0.5) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black') + ylim(0, 3.2)

ggplot(g, aes(visibility, dewlappermin)) +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = res_dewsize), alpha = 0.5) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black') + ylim(0, 3.2)



# lineatopus
ggplot(l, aes(visibility, dewlappermin)) +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = dewsizemax), alpha = 0.5) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black', linetype = 2) 

ggplot(l, aes(visibility, dewlappermin)) +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = res_dewsize), alpha = 0.5) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black', linetype = 2) 

ggplot(l, aes(dewsizemax, dewlappermin)) +
  #geom_point(aes(colour = visibility)) + 
  geom_point(aes(size = dewsizemax), alpha = 0.5) + scale_radius(range = c(2, 12)) +
  theme_classic() +
   geom_smooth(method = 'lm', se = FALSE, color = 'black') 

```






###########################3
Function to run regression models
```{r regression models grahami}
library(tidyverse)
library(knitr)
library(broom)

# creating a funciton to run the regression analyses 
lmfxn <- function(response, predictors, speciesname){
  
   fixedformula <- paste(response, "~", predictors)
   
   lm(as.formula(fixedformula), data = subset(jmanoles, species == speciesname)) %>%
     tidy() %>%
     kable()

}


```
**Full Model (absolute dewlap size)**   
`dewsizemax` + `visibility`
```{r full model absolute}

lmfxn('dewlappermin', 'dewsizemax + visibility', 'grahami')

lmfxn('dewlappermin', 'dewsizemax + visibility', 'lineatopus')

lmfxn('dewlappermin', 'dewsizemax + visibility', 'valencienni')

```

**Full Model with Intereaction**
```{r full model absolute interaction}

lmfxn('dewlappermin', 'dewsizemax * visibility', 'grahami')

lmfxn('dewlappermin', 'dewsizemax * visibility', 'lineatopus')


```

**Full Model (relative dewlap size)**   
`svl` + `dewsizemax` + `visibility`
```{r full model relative}

lmfxn('dewlappermin', 'svl + dewsizemax + visibility', 'grahami')

lmfxn('dewlappermin', 'svl + dewsizemax + visibility', 'lineatopus')

lmfxn('dewlappermin', 'svl + dewsizemax + visibility', 'valencienni')

```


**Signal Size Model**  
Do males with larger dewlaps (absolute size) display their dewlaps more than lizards with smaller dewlaps?  
`dewsizemax`  
  
Do males with relatively larger dewlaps display their dewlaps more than lizards with smaller dewlaps?
`dewsizemax` + `svl`
```{r signal size model}

lmfxn('dewlappermin', 'dewsizemax', 'grahami')
lmfxn('dewlappermin', 'dewsizemax + svl', 'grahami')
lmfxn('dewlappermin', 'svl', 'grahami')


lmfxn('dewlappermin', 'dewsizemax', 'lineatopus')
lmfxn('dewlappermin', 'dewsizemax + svl', 'lineatopus')
lmfxn('dewlappermin', 'svl', 'lineatopus')


lmfxn('dewlappermin', 'dewsizemax', 'valencienni')
lmfxn('dewlappermin', 'dewsizemax + svl', 'valencienni')
lmfxn('dewlappermin', 'svl', 'valencienni')

```
```{r}
# library(ppcor)
# 
# cor <- data.frame(g$dewlappermin, g$dewsizemax, g$svl)
# spcor.test(g$dewlappermin, g$dewsizemax, g$svl, method = c('pearson'))

```


```{r fig svl dewsizemax}
library(ggplot2)

ggplt <- ggplot(jmanoles,aes(x = svl, y = dewsizemax, shape = sex, color = species))+
         geom_point()+
         theme_classic() 

ggplt+geom_smooth(method=lm,se=FALSE,fullrange=TRUE,
                  aes(color=species))

```

```{r fig svl dewlappermin}
library(ggplot2)

ggplt <- ggplot(jmanoles,aes(x = svl, y = dewlappermin, shape = species, color = species))+
         geom_point()+
         theme_classic() + ylim(0,3.2)

ggplt+geom_smooth(method=lm,se=FALSE,fullrange=TRUE,
                  aes(color=species))

```

```{r fig dewsizemax dewlappermin}
library(ggplot2)

ggplt <- ggplot(jmanoles,aes(x = dewsizemax, y = dewlappermin, shape = species, color = species))+
         geom_point()+
         theme_classic() + ylim(0,3.2)

ggplt+geom_smooth(method=lm,se=FALSE,fullrange=TRUE,
                  aes(color=species))

```

```{r fig dewsizemax/svl dewlappermin}
library(ggplot2)

ggplt <- ggplot(jmanoles,aes(x = dewsizemax/svl, y = dewlappermin, shape = species, color = species))+
         geom_point()+
         theme_classic() + ylim(0,3.2)

ggplt+geom_smooth(method=lm,se=FALSE,fullrange=TRUE,
                  aes(color=species))

```


**Visibility Model**   
Do males with in more visible habitats display their dewlap more than lizards in more cluttered habitats?  
`visibility`  
  
```{r visibility model}

lmfxn('dewlappermin', 'visibility', 'grahami')

lmfxn('dewlappermin', 'visibility', 'lineatopus')

lmfxn('dewlappermin', 'visibility', 'valencienni')

```

```{r fig visibility dewlapppermin}
library(ggplot2)

#ggplot(data = jmanoles, aes(x = visibility, y = dewlappermin, color = species)) + geom_point()

ggplt <- ggplot(jmanoles,aes(x = visibility, y = dewlappermin, shape = species, color = species))+
         geom_point()+
         theme_classic() + ylim(0,3.2) + xlim(0.59,1)

ggplt+geom_smooth(method=lm,se=FALSE,fullrange=TRUE,
                  aes(color=species))

```

Note: this relationship, especially for grahami seems more curvilinear than linear
