---
title: "Repeatability Dewlap Behavior Analysis"
author: "Lauren E Johnson"
date: "12/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Import data from prior analyses
```{r}
load(here::here('data/jm_anoles.Rda'))

#Creating separate dataframes for each species
g <- jmanoles[which(jmanoles$species == "grahami"),]
l <- jmanoles[which(jmanoles$species == "lineatopus"),]
v <- jmanoles[which(jmanoles$species == "valencienni"),]
```

Import raw unorganized data wrangle behavior data
```{r}
# Read data file
gra_beh <- read.csv(here::here('data_raw/GRA_beh.csv'))
lin_beh <- read.csv(here::here('data_raw/LIN_beh.csv'))

# Create column with dewlap rate per minute, since not all observations were exactly 20 minutes 
gra_beh$dewlappermin <- ((gra_beh$Dewlap)/(gra_beh$Total.time))
lin_beh$dewlappermin <- ((lin_beh$Dewlap)/(lin_beh$Totaltime))


# Numbering rows for each data frame individual
library(dplyr)
gra_beh <- gra_beh %>% group_by(ID) %>% mutate(observation = row_number())
lin_beh <- lin_beh %>% group_by(ID) %>% mutate(observation = row_number())


# Creating a single wide row for each individual
library(tidyr)
gra_beh_wide <- pivot_wider(data = gra_beh, names_from = observation, 
                            names_glue = "{observation}_{.value}", 
                            values_from = c(Date, Person, Start.time, Total.time,
                                            Dewlap, Pushbobs, Movements))

lin_beh_wide <- pivot_wider(data = lin_beh, names_from = observation,
                              names_glue = "{observation}_{.value}",
                              values_from = c(Date, Person, Starttime, Totaltime,
                                              Dewlap, Pushbobs, Movements))
```

Import raw unorganized data and data wrangle habitat data
```{r}
# Read data file
gra_hab <- read.csv(here::here('data_raw/GRA_hab.csv'))
lin_hab <- read.csv(here::here('data_raw/LIN_hab.csv'))

# Remove rows that do not contain visibility data
gra_hab <- gra_hab[!is.na(gra_hab$Vis),]
lin_hab <- lin_hab[!is.na(lin_hab$Vis),]


# Numbering rows for each data frame individual
library(dplyr)
gra_hab <- gra_hab %>% group_by(ID) %>% mutate(observation = row_number())
lin_hab <- lin_hab %>% group_by(ID) %>% mutate(observation = row_number())

# Creating a single wide row for each individual
library(tidyr)
gra_hab_wide <- pivot_wider(data = gra_hab,
                              names_from = observation,
                              names_glue = "{observation}_{.value}",
                              values_from = c(height, diameter, Vis))
lin_hab_wide <- pivot_wider(data = lin_hab,
                              names_from = observation,
                              names_glue = "{observation}_{.value}",
                              values_from = c(height, diameter, Vis))

```

Create dataframe for grahami with averages and standard deviations for behavior and habitat variables
```{r}
library(tidyverse)
# grahami 

# Overall total dewlaps per total time 
dew_rate_df <- gra_beh %>%
  group_by(ID) %>%
  summarise(dew_rate = (sum(Dewlap)/sum(Total.time)))

# Average dewlap rate across all observations 
dew_rate_mean_df <- gra_beh %>%
  group_by(ID) %>%
  summarise(dew_rate_mean = mean(dewlappermin))

# Standard deviation of average dewlap rate across all observations
dew_rate_sd_df <- gra_beh %>%
  group_by(ID) %>%
  summarise(dew_rate_sd = sd(dewlappermin))

# Average visibility of perch across all observations
vis_mean_df <- gra_hab %>% 
  group_by(ID) %>%
  summarise(vis_mean = mean(Vis, na.rm = T))

# Standard deviation of perch visibility across all observations
vis_sd_df <- gra_hab %>% 
  group_by(ID) %>%
  summarise(vis_sd = sd(Vis, na.rm = T))

# Number of behavioral observations for each individual 
dew_rate_observations_df <- gra_beh %>%
  count(ID, name = 'dewlappermin_observations')

# Number of habitat observations for each individual 
vis_observations_df <- gra_hab %>% 
  count(ID, name = 'vis_observations')


# Create a list of all the summary information dataframes 
df_list <- list(dew_rate_df, dew_rate_mean_df, dew_rate_sd_df, vis_mean_df, vis_sd_df, dew_rate_observations_df, vis_observations_df, g)

# Join all the dataframe into a composite dataframe from wich we can run analyses etc. 
gra_df <- Reduce(function(d1, d2) merge(d1, d2, by = 'ID', all.x = T), df_list)


```

Create dataframe for lineatopus with averages and standard deviations for behavior and habitat variables
```{r}
library(tidyverse)
# grahami 

# Overall total dewlaps per total time 
dew_rate_df <- lin_beh %>%
  group_by(ID) %>%
  summarise(dew_rate = (sum(Dewlap)/sum(Totaltime)))

# Average dewlap rate across all observations 
dew_rate_mean_df <- lin_beh %>%
  group_by(ID) %>%
  summarise(dew_rate_mean = mean(dewlappermin))

# Standard deviation of average dewlap rate across all observations
dew_rate_sd_df <- lin_beh %>%
  group_by(ID) %>%
  summarise(dew_rate_sd = sd(dewlappermin))

# Average visibility of perch across all observations
vis_mean_df <- lin_hab %>% 
  group_by(ID) %>%
  summarise(vis_mean = mean(Vis, na.rm = T))

# Standard deviation of perch visibility across all observations
vis_sd_df <- lin_hab %>% 
  group_by(ID) %>%
  summarise(vis_sd = sd(Vis, na.rm = T))

# Number of behavioral observations for each individual 
dew_rate_observations_df <- lin_beh %>%
  count(ID, name = 'dewlappermin_observations')

# Number of habitat observations for each individual 
vis_observations_df <- lin_hab %>% 
  count(ID, name = 'vis_observations')


# Create a list of all the summary information dataframes 
df_list <- list(dew_rate_df, dew_rate_mean_df, dew_rate_sd_df, vis_mean_df, vis_sd_df, dew_rate_observations_df, vis_observations_df, l)

# Join all the dataframe into a composite dataframe from wich we can run analyses etc. 
lin_df <- Reduce(function(d1, d2) merge(d1, d2, by = 'ID', all.x = T), df_list)


```

mean and standard deviation for number of observations for behavior and habitat selection 
```{r}
summary(lin_df$dewlappermin_observations)
mean(lin_df$dewlappermin_observations)
sd(lin_df$dewlappermin_observations)


summary(gra_df$dewlappermin_observations)
mean(gra_df$dewlappermin_observations)
sd(gra_df$dewlappermin_observations)


summary(lin_df$vis_observations)
mean(lin_df$vis_observations, na.rm = T)
sd(lin_df$vis_observations, na.rm = T)


summary(gra_df$vis_observations)
mean(gra_df$vis_observations, na.rm = T)
sd(gra_df$vis_observations, na.rm = T)
```


Grahami Plots
```{r}

# This graph is the keeper because it scales it to log10 but keeps the axis with the true lables not the weird lables that don't mean anything intuitively 
ggplot(gra_df, aes(vis_mean, dew_rate)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
   geom_smooth(method = lm, se = FALSE)

ggplot(gra_df, aes(svl, dew_rate)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

ggplot(gra_df, aes(dewsizemax, dew_rate)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

# residual dewlap plot is in the dewlap_analysis folder



```

Lineatopus Plots
```{r}

# This graph is the keeper because it scales it to log10 but keeps the axis with the true lables not the weird lables that don't mean anything intuitively 
ggplot(lin_df, aes(vis_mean, dew_rate)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
   geom_smooth(method = lm, se = FALSE)

ggplot(lin_df, aes(svl, dew_rate)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

ggplot(lin_df, aes(dewsizemax, dew_rate)) +
  geom_point() + 
  theme_classic() +
  scale_y_log10() +
  scale_x_log10()+
   geom_smooth(method = lm, se = FALSE)

```

Analyses
Grahami - visibility and dewsize max are associated with dewlap display
Lineatopus - only dewsize max is associated with dewlap display 
for neither species is SVL associated with dewlap display (there are some differences in svl and dewlap size)
```{r}
dew.lm <- lm(log(dew_rate + 1) ~ visibility * dewsizemax, data = gra_df)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm)

dew.lm <- lm(log(dew_rate + 1) ~ visibility + dewsizemax, data = gra_df)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm)

dew.lm <- lm(log(dew_rate + 1) ~ visibility * svl, data = gra_df)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm) 

dew.lm <- lm(log(dew_rate + 1) ~ visibility + svl, data = gra_df)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm) 


dew.lm <- lm(log(dew_rate + 1) ~ visibility * dewsizemax, data = lin_df)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm) 

dew.lm <- lm(log(dew_rate + 1) ~ visibility + dewsizemax, data = lin_df)
plot(dew.lm)
summary(dew.lm)
car::Anova(dew.lm) 

dew.lm <- lm(log(dew_rate + 1) ~ visibility * svl, data = lin_df)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 

dew.lm <- lm(log(dew_rate + 1) ~ visibility + svl, data = lin_df)
plot(dew.lm)
summary(dew.lm) 
car::Anova(dew.lm) 
```

Create lineplot showing intra-individual variation in dewlap display and habitat characteristics
```{r}
lin_hab_2<-lin_hab[!(lin_hab$ID=='L109' | lin_hab$ID=='L124' | lin_hab$ID=='L150' | lin_hab$ID=='L160' | lin_hab$ID== 'L166' | lin_hab$ID== 'L167' | lin_hab$ID== 'L169' | lin_hab$ID== 'L176' | lin_hab$ID== 'L305' | lin_hab$ID== 'L321' | lin_hab$ID== 'L323' | lin_hab$ID== 'L324' | lin_hab$ID== 'L331' | lin_hab$ID== 'L334'),]

gra_hab_2 <- gra_hab[!(gra_hab$ID=='G107' | gra_hab$ID=='G120' | gra_hab$ID=='G123' | gra_hab$ID=='G127' | gra_hab$ID=='G128' | gra_hab$ID=='G137' | gra_hab$ID=='G142' | gra_hab$ID=='G150' | gra_hab$ID=='G159' | gra_hab$ID=='G305' | gra_hab$ID=='G334' | gra_hab$ID=='G338'),]


library(ggplot2)


library(tidyverse)
dMean <- gra_beh %>%
    group_by(ID) %>%
    summarise(MN = mean(dewlappermin))



# Repeatability of dewlappermin for grahami 
grahbehMean <- gra_beh %>%
    group_by(ID) %>%
    summarise(MN = mean(dewlappermin))

ggplot(data = gra_beh, aes(x = observation, y = dewlappermin, group = ID)) +
  geom_hline(data = grahbehMean, aes(yintercept = MN)) +
  geom_line(aes(color = ID)) +
  geom_point(aes(color = ID)) +
  facet_wrap(. ~ ID, ncol = 10) +
  theme(legend.position = 'none')

# Repeatability of dewlappermin for lineatopus 
linbehMean <- lin_beh %>%
    group_by(ID) %>%
    summarise(MN = mean(dewlappermin))

ggplot(data = lin_beh, aes(x = observation, y = dewlappermin, group = ID)) +
  geom_hline(data = linbehMean, aes(yintercept = MN)) +
  geom_line(aes(color = ID)) +
  geom_point(aes(color = ID)) +
  facet_wrap(. ~ ID, ncol = 10) +
  theme(legend.position = 'none')

# Repeatability of perch visibility for grahami 

grahVisMean <- gra_hab_2 %>%
    group_by(ID) %>%
    summarise(MN = mean(Vis))

ggplot(data = gra_hab_2, aes(x = observation, y = Vis, group = ID)) +
  geom_hline(data = grahVisMean, aes(yintercept = MN)) +
  geom_path(aes(color = ID)) +
  geom_point(aes(color = ID)) +
  facet_wrap(. ~ ID, ncol = 10) +
  theme(legend.position = 'none')


# Repeatability of perch visibility for lineatopus 

linVisMean <- lin_hab_2 %>%
    group_by(ID) %>%
    summarise(MN = mean(Vis))

ggplot(data = lin_hab_2, aes(x = observation, y = Vis, group = ID)) +
  geom_hline(data = linVisMean, aes(yintercept = MN)) +
  geom_line(aes(color = ID)) +
  geom_point(aes(color = ID)) +
  facet_wrap(. ~ ID, ncol = 10) +
  theme(legend.position = 'none')

# Perch height
grahheightMean <- gra_hab_2 %>%
    group_by(ID) %>%
    summarise(MN = mean(height))

ggplot(data = gra_hab_2, aes(x = observation, y = height, group = ID)) +
  geom_hline(data = grahheightMean, aes(yintercept = MN)) +
  geom_line(aes(color = ID)) +
  geom_point(aes(color = ID)) +
  facet_wrap(. ~ ID, ncol = 10) +
  theme(legend.position = 'none')


linheightMean <- lin_hab_2 %>%
    group_by(ID) %>%
    summarise(MN = mean(height))

ggplot(data = lin_hab_2, aes(x = observation, y = height, group = ID)) +
  geom_hline(data = linheightMean, aes(yintercept = MN)) +
  geom_line(aes(color = ID)) +
  geom_point(aes(color = ID)) +
  facet_wrap(. ~ ID, ncol = 10) +
  theme(legend.position = 'none')

# # Perch diameter - issue in that some are just written as rock of wall...
# ggplot(data = gra_hab, aes(x = observation, y = diameter, group = ID)) +
#   geom_path(aes(color = ID)) +
#   geom_point(aes(color = ID)) +
#   facet_wrap(. ~ ID, ncol = 10) +
#   theme(legend.position = 'none')
# 
# ggplot(data = lin_hab, aes(x = observation, y = diameter, group = ID)) +
#   geom_line(aes(color = ID)) +
#   geom_point(aes(color = ID)) +
#   facet_wrap(. ~ ID, ncol = 10) +
#   theme(legend.position = 'none')
```


Rpeateability analyeses
```{r}

library(rptR)

rpt(Vis ~ (1 | ID), grname = "Visibility", data = lin_hab, link = 'logit', datatype = "Proportion", 
    nboot = 0, npermut = 0)

```

